---
title: "ST 558 Homework 4"
author: "Luke Freudenheim"
description: In this report we will first answer conceptual questions about certain functions within BaseR and dplyr. We will then move into function creation. We will create functions that will calculate both the MAE and RMSE of given response values and prediction values. Lastly, we will query to an API by first entering a specific URL and eventually create a function that automates this process given an API key, a title, and a date.
format: html
---

# Task 1: Conceptional Questions

1. What is the purpose of the lapply() function? What is the equivalent purrr function?

> This function take in either a data frame, vector or list as input and always outputs a list after applying a specific function to your data. The equivalent purrr function is the map() function.

2. Suppose we have a list called my_list. Each element of the list is a numeric data frame (all columns
are numeric). We want use lapply() to run the code cor(numeric_matrix, method = "kendall")
on each element of the list. Write code to do this below!

> lapply(numeric_matrix = my_list, FUN = cor, method = "kendall")

3. What are two advantages of using purrr functions instead of the BaseR apply family?

> The main advantage of the purrr functions is there is more consistency within the function family than there is with the BaseR family. The other advantage is that purrr functions give a shorthand way to make anonymous functions, which is helpful.

4. What is a side-effect function?

> A side effect funciton is a function that doesn't try to change the data you are using, but instead creates output of your data without altering it.

5. Why can you name a variable sd in a function and not cause any issues with the sd function?

> You can name a variable sd within a function and it won't cause issues with the sd function because within functions a temporary environment is created that is separate from the global environment. Therefore, the temporary environment will have an sd variable stored in it that is not stored in the global environment.

# Task 2: Writing R Functions

## Loading necessary packages

```{r}
library(tidyverse)
library(Metrics)
```


## Function that Returns RMSE

This function will take in two vectors (a response vector and a prediction vector) and will compute and return the RMSE of the vectors.

```{r}
getRMSE <- function(response, prediction, ...) {
  errors <- (response - prediction)^2
  rmse <- sqrt(mean(errors, ...))
  return(rmse)
}
```

## Testing our Function with Random Data

Here we will test our getRMSE function we created on data that is randomized in order to see if it runs properly.

```{r}
# Generating random data
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
# Testing Our Function
getRMSE(resp, pred)
# Confirming this is the correct RMSE using premade function
rmse(resp, pred)
```

Next, we will replace two of the response values with missing values, and then test our function with and without removing the NA values. We will first test the function without removing NA values here:

```{r}
# Replacing response values with missing values
resp[1] <- resp[2] <- NA_real_
# Testing function without removing NA values
getRMSE(resp, pred)
```

Here we will specify the na.rm argument and test the function again:

```{r}
# Testing function with removing NA values
getRMSE(resp, pred, na.rm = TRUE)
```

## Function for MAE

In this code, we will create a function that also takes in a vector of response values and a vector of prediction values. This function will calculate the MAE of the vectors given.

```{r}
getMAE <- function(response, prediction, ...) {
  errors <- abs(response - prediction)
  mae <- mean(errors, ...)
  return(mae)
}
```

## Testing our Function with Random Data

We will now generate random response and prediction data to test the getMAE function we made.

```{r}
# Generating random data
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
# Testing our getMAE function
getMAE(resp, pred)
# Confirming this is the correct MAE using premade function
mae(resp, pred)
```

Next, we will replace two of the response values with missing values, and then test our function with and without removing the NA values. We will first test the function without removing NA values here:

```{r}
# Replacing response values with missing values
resp[1] <- resp[2] <- NA_real_
# Testing function without removing NA values
getMAE(resp, pred)
```

Here we will specify the na.rm argument and test the function again:

```{r}
# Testing function with removing NA values
getMAE(resp, pred, na.rm = TRUE)
```

## Wrapper Function Creation

```{r}
metric_wrapper <- function(response, prediction, metric = c("RMSE", "MAE"), ...) {
  if(!(is.vector(response) & is.vector(prediction) & is.atomic(response) & is.atomic(prediction) & is.numeric(response) & is.numeric(prediction))) {
    stop("One or both inputs are not numeric atomic vectors.")
  }
  else if("RMSE" %in% metric & "MAE" %in% metric) {
    rmse <- getRMSE(response, prediction, ...)
    mae <- getMAE(response, prediction, ...)
    return(list("Root Mean Square Error" = rmse, "Mean Absolute Error" = mae))
  }

  else if("MAE" %in% metric) {
    mae <- getMAE(response, prediction, ...)
    return(list("Mean Absolute Error" = mae))
  } else if("RMSE" %in% metric) {
    rmse <- getRMSE(response, prediction, ...)
    return(list("Root Mean Squared Error" = rmse))
  } else {
    stop("Wrong metric specified.")
  }
}
```

## Testing our Wrapper Function

Next we will test our wrapper function on randomly generated response and prediction data. For the first test we will specify both metrics.

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10*x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
# Calling our function specifying both metrics
metric_wrapper(resp, pred, c("MAE", "RMSE"))
```

Next, we will test our function by only specifying RMSE as the metric.

```{r}
metric_wrapper(resp, pred, "RMSE")
```

Next we will test our function by just specifying MAE as the metric.

```{r}
metric_wrapper(resp, pred, "MAE")
```

Next, we will test our function with both metrics and each metric separately after putting two missing values into the response variables. Here, we will test the function without specifying the na.rm function.

```{r}
# Adding the NA Values
resp[1] <- resp[2] <- NA_real_
# Testing the function with both metrics
metric_wrapper(resp, pred, c("RMSE", "MAE"))
# Testing the function specifying RMSE
metric_wrapper(resp, pred, "RMSE")
# Testing the function specifying MAE
metric_wrapper(resp, pred, "MAE")
```

Now we will specify the na.rm = TRUE argument within our function calls and repeat with both metrics and each one separately.

```{r}
# Testing the function with both metrics
metric_wrapper(resp, pred, c("RMSE", "MAE"), na.rm = TRUE)
# Testing the function specifying RMSE
metric_wrapper(resp, pred, "RMSE", na.rm = TRUE)
# Testing the function specifying MAE
metric_wrapper(resp, pred, "MAE", na.rm = TRUE)
```

Next, we will pass our wrapper function a data frame instead of a vector, to see if the function stops and an error message occurs.

```{r, error=TRUE}
resp <- data.frame(resp)
metric_wrapper(response = resp, pred)
```

Our function successfully stopped and sent out the correct error message when passing it a data frame, which means it is working correctly.


# Task 3: Querying an API and a Tidy-Style Function

We will now query an API and obtain a data set with the top headlines from TechCrunch in recent months.

```{r}
library(httr)
library(jsonlite)
URL <- "https://newsapi.org/v2/top-headlines?sources=techcrunch&apiKey=e04b926ae2dc4c96a39f4d4011adf87a"
tech_crunch <- httr::GET(URL)
parsed_tech_crunch_data <- fromJSON(rawToChar(tech_crunch$content))
final_tech_crunch <- as_tibble(parsed_tech_crunch_data$articles)
final_tech_crunch
```

Next, we will write a function that takes in the title or subject to search for, the time period to search from, and the API key. It will output the data frame of the articles in that subject and time period.

```{r}
API_Query <- function(title, time, api_key) {
url <- paste("https://newsapi.org/v2/everything?q=", title, "&from=", time, "&apiKey=", api_key, sep = "")
raw_data <- httr::GET(url)
parsed_data <- fromJSON(rawToChar(raw_data$content))
final_data <- as_tibble(parsed_data$articles)
return(final_data)
}
```

Here, we will test our API_Query function that we just made using two different dates and subjects. The first title we will use is tesla, so we will try to find the articles from the last month regarding tesla.

```{r}
API_Query("tesla", "2024-05-26", "e04b926ae2dc4c96a39f4d4011adf87a")
```

Our function was successful in returning the data frame with articles containing tesla from the last month.

Next, we will use the subject "sports" and we will change the date given. We will see if this successfully returns the articles regarding sports from the date given until now.

```{r}
API_Query("sports", "06-20-2024", "e04b926ae2dc4c96a39f4d4011adf87a")
```

Our function successfully allows us to query this API!
